Insilico Splits

### 2018-05-17 AllSamplesModel.Rmd
...

cache.dir = paste(data, 'Analysis/5e6_arms_all_exAHM0320', sep='/')

@@

p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
  geom_text( aes(label=round(mean, 2)), vjust=2, size=5, color='gray39', show.legend=F) +
  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
#  geom_point(size=2) +
  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
  ylim(0,1)
p

#ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```

@@

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme
p
ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)

# oddPt = names(pg.samp)[which.max(performance.at.1se)]
# subset(sum.patient.data, Hospital.Research.ID == oddPt)
# pg.samp[[oddPt]]$Prediction
# 
# oddPt = names(pg.samp)[which.min(performance.at.1se)]

# Not the pt with the most samples or endos
#sum.patient.data[which.max(sum.patient.data$total.samples),]
```

```{r echo=F, eval=F}
## All LOO
files = list.files('~/Data/Ellie/Analysis/5e6_arms_LOO', full.names = T)

sum.patient.data$Hospital.Research.ID[1]

pg.samp = list()
performance = vector(length=nrow(sum.patient.data), mode='numeric')
names(performance) = sum.patient.data$Hospital.Research.ID

for (pt in sum.patient.data$Hospital.Research.ID) {
  ptF = grep(pt, files, value=F)
  load(grep('loo\\.Rdata', ptF, value=T), verbose=T)
  #load(grep('model_data\\.Rdata', ptF, value=T), verbose=T)
  performance[pt] = performance.at.1se
  pg.samp[[pt]] = info
}
rm(plots,coefs,nzcoefs,fits,info)

#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_loo.Rdata", verbose=T)
#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/PR1_HIN_042_model_data.Rdata", verbose=T)

#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_loo.Rdata", verbose=T)
#load("/Users/skillcoy/Data/Ellie/Analysis/5e6_arms_LOO/AHM0320_model_data.Rdata", verbose=T)

fitLOO <- glmnet(dysplasia.df, labels[rownames(dysplasia.df)], alpha=0.9, family='binomial', nlambda=500, standardize=F)

pm = predict(fitLOO, newx=allDf, s=fits$lambda.1se, type='response')
or = predict(fitLOO, newx=allDf, s=fits$lambda.1se, type='link')

hist(pm)
hist(or)

pm[subset(patient.info, Hospital.Research.ID == 'AHM0320')$Samplename,]
or[subset(patient.info, Hospital.Research.ID == 'AHM0320')$Samplename,]

sort(performance)

pg.samp[which.max(performance)]

pg.samp[which.min(performance)]

back = do.call(rbind, pg.samp)

subset(sum.patient.data, Status == 'NP' & total.endos <=2)

pm[pg.samp[['AD0309']]$Samplename,]
pg.samp[['AD0309']]$Prediction

@@

roc$is = 'all'

@@

load(file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'), verbose=T)
insilico.rocs[['all']] = roc

@@

rr = ggplot(subset(df, is == 'all' | auc %in% c(range(auc))), aes(Specificity, Sensitivity)) + 
  geom_line(aes(color=factor(model)), size=1) + 
  geom_line(data=subset(df, auc == min(auc)), color='darkblue', linetype='dotted') + 
  geom_line(data=subset(df, auc == max(auc)), color='darkblue', linetype='dotted') +
  scale_x_reverse() + scale_color_brewer(palette='Paired') + 
  geom_label(data=subset(best, model == 'all' | auc %in% c(min(auc), max(auc)) ), aes(specificity, sensitivity, label=paste('AUC ',round(auc, 2), sep='\n'))) +
  labs(title='ROC CI', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
  theme(legend.position = 'none')
rr  
ggsave(filename='plots/auc/roc_CI.png', plot=rr, width=4, height=4, units='in', dpi=300, scale=1.5)

@@

ft.fun<-function(NP,P) {
  f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  cbind.data.frame('p.value'=f$p.value)
  #cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

qt = cbind(qt, back %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))#, 
pred.confidence = qt %>% as.data.frame.matrix %>% group_by(quant) %>% dplyr::mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value,  'conf'=ifelse(p.value < 0.05, '*', '') )

@@

pred.confidence$Risk = 'Moderate'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .5)) ] = 'Low'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

@@

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
#pred.confidence$Risk[1:5] = 'Low'
#pred.confidence$Risk[6:10] = 'High'
pred.confidence$Risk[4:7] = 'Moderate'

@@

## count arms differently than segs??
df2 = df %>% group_by(Status, Patient, Endoscopy.Year) %>% summarise( 
  'max.Pred'=max(Prediction),
  'max.RR'=max(OR),
  'max.Path'=max(Pathology),
  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
  )
#as.data.frame(df2)
#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))

df2 = df2 %>% group_by(Status, Patient) %>% mutate(
  'early'= Endoscopy.Year == min(Endoscopy.Year),
  'late' = Endoscopy.Year == max(Endoscopy.Year)
)
df2 = subset(df2, Patient %in% names(which(table(df2$Patient) > 1)))

el = subset(df2, early | late)
el2 = el %>% group_by(Status, Patient) %>% summarise( diff(arms), diff(CNA) )
wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
if (wtA < 0.0001) wtA = '< 0.0001'
#wilcox.test(subset(el2, Status == 'Progressor')$`diff(CNA)`, subset(el2, Status == 'Non-Progressor')$`diff(CNA)`)

@@

second = subset(rules,  t2 != 'Endpoint') %>% 
  group_by(Patient, Status) %>% 
  summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

@@

with(second, table(Status, rule))
```

@@

```{r 5yrrisk, echo=F, fig.height=8, fig.width=12}
cases = table(sum.patient.data$Status)['P']

>>>>

cache.dir = paste(data, 'Analysis/5e6_arms_all_logR', sep='/')

@@

p = ggplot(classification.rate, aes(model, mean, group=model, fill=model)) + geom_col() +
  geom_text( aes(label=round(mean, 3)), vjust=2, size=5, color='gray39', show.legend=F) +
  geom_text( aes(y=0.6, label=paste(all.coef, " coef\n(", stable.coef, ")", sep='') ), color='gray39', size=5) +
  geom_errorbar(aes(ymin=mean-sme, ymax=mean+sme),color='grey39', width=0.1, size=1) + 
#  geom_point(size=2) +
  labs(x='', y='Classification rate', title=paste('Classification rate per model at', select.alpha)) + 
  plot.theme + theme(axis.text.x=element_text(angle=45,hjust=1,face="bold",size=12), legend.position = 'none') +
  ylim(0,1)
p

ggsave('plots/performance/compare_classification.png', plot=p, height=5, width=3, units='in', dpi=300 )
```

@@

p = ggplot(df, aes(y=performance.at.1se, x='LOO')) + 
  geom_boxplot(fill='lightblue', color='darkgrey', outlier.fill=NA, outlier.color=NA, size=0.8) + geom_jitter(color='grey39', width=0.3) +
  geom_label(data=subset(df, performance.at.1se %in% range(performance.at.1se)), aes(y=performance.at.1se, label=round(performance.at.1se, 3)), fontface='bold', fill='lightblue') +
  labs(y='Classification rate', x='LOO model', title='Performance for LOO', subtitle='All samples model') +
  plot.theme
p
ggsave('plots/performance_loo.png', plot=p, width=5, height=5, units='in', dpi=300)
```

@@

roc$is = 'all'
roc.plot(roc)

@@

insilico.rocs = list()
ifile = paste(cache.dir, 'insilico_splits.Rdata', sep='/')
if (file.exists(ifile)) {
  load(ifile, verbose=T)
  insilico.rocs[['all']] = roc

@@

  rr = ggplot(subset(df, is == 'all' | auc %in% c(range(auc))), aes(Specificity, Sensitivity)) + 
    geom_line(aes(color=factor(model)), size=1) + 
    geom_line(data=subset(df, auc == min(auc)), color='darkblue', linetype='dotted') + 
    geom_line(data=subset(df, auc == max(auc)), color='darkblue', linetype='dotted') +
    scale_x_reverse() + scale_color_brewer(palette='Paired') + 
    geom_label(data=subset(best, model == 'all' | auc %in% c(min(auc), max(auc)) ), aes(specificity, sensitivity, label=paste('AUC ',round(auc, 2), sep='\n'))) +
    labs(title='ROC CI', x='Specificity (1-FPR)', y='Sensitivity (TPR)')  + plot.theme +
    theme(legend.position = 'none')
  rr  
  ggsave(filename='plots/auc/roc_CI.png', plot=rr, width=6, height=6, units='in', dpi=300, scale=1.5)
} else {
  message("No insilico split file.")
}

@@

ft.fun<-function(NP,P) {
  #f = chisq.test(rbind(cbind(NP,P),table(sum.patient.data$Status)))
  f = chisq.test(c(NP,P))
  cbind.data.frame('p.value'=f$p.value)
  #cbind.data.frame('p.value'=f$p.value, 'odds.ratio'=f$estimate)
}

qt = cbind(qt, back %>% group_by(quants) %>% dplyr::summarise ( 'mn'=mean(Prediction), 'sd'=sd(Prediction) ))
pred.confidence = qt %>% as.data.frame.matrix %>% rowwise() %>% dplyr::mutate( 
  'perc'=P/sum(NP,P), 'p.value'=ft.fun(NP,P)$p.value, 'conf'=ifelse(p.value < 0.05, '*', '') )

@@

pred.confidence$Risk = 'Moderate'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc < .3)) ] = 'Low'
pred.confidence$Risk[ which(with(pred.confidence, p.value < 0.05 & perc > .5)) ] = 'High'

@@

```{r pcalib, echo=F, warning=F, message=F, fig.width=8, fig.height=6}
#pred.confidence$Risk[1:5] = 'Low'
#pred.confidence$Risk[6:10] = 'High'
#pred.confidence$Risk[4:7] = 'Moderate'

@@

## count arms differently than segs??
df2 = df %>% group_by(Status, Patient, Endoscopy.Year) %>% dplyr::summarise( 
  'max.Pred'=max(Prediction),
  'max.RR'=max(OR),
  'max.Path'=max(Pathology),
  'CNA'=length(which(seg.type != 'arms' & cn %in% c('loss','gain'))),
  'arms'=length(which(seg.type == 'arms' & cn %in% c('loss','gain'))),
  'ratio'=length(which(seg.type != 'arms' & cn %in% c('loss','gain')))/length(cn)
  )
#as.data.frame(df2)
#df2 = subset(df2, max.Path %in% c('BE','ID','LGD'))

df2 = df2 %>% group_by(Status, Patient) %>% dplyr::mutate(
  'early'= Endoscopy.Year == min(Endoscopy.Year),
  'late' = Endoscopy.Year == max(Endoscopy.Year)
)
df2 = subset(df2, Patient %in% names(which(table(df2$Patient) > 1)))

el = subset(df2, early | late)
el2 = el %>% group_by(Status, Patient) %>% dplyr::summarise( diff(arms), diff(CNA) )
wtA = round(wilcox.test(subset(el2, Status == 'Progressor')$`diff(arms)`, subset(el2, Status == 'Non-Progressor')$`diff(arms)`)$p.value, 3)
if (wtA < 0.0001) wtA = '< 0.0001'
#wilcox.test(subset(el2, Status == 'Progressor')$`diff(CNA)`, subset(el2, Status == 'Non-Progressor')$`diff(CNA)`)

@@

second = subset(rules,  t2 != 'Endpoint') %>% 
  group_by(Patient, Status) %>% 
  dplyr::summarise(  max(t2), 'rule'=rule[t2 == max(t2)]  )

@@

@@

```{r 5yrrisk, echo=F, fig.height=8, fig.width=12, eval=F}
cases = table(sum.patient.data$Status)['P']

### 2018-05-17 Insilico Splits.Rmd
...

data = '~/Data/Ellie'
tile.w=c(5e06, 'arms')

includeDemo = F

@@

load(grep('All_patients.Rdata', analysis.files, value=T), verbose=T)

@@

#sum.pt = summarise.patient.info(all.patient.info)

sum.pt = all.patient.info %>% group_by(Set, Patient) %>% summarize(
  med.reads=median(Total.Reads),
  sd.reads=sd(Total.Reads),
  n.endoscopy=length(unique(PID)),
  n.years=unique(Final.Endoscopy-Initial.Endoscopy),
  n.samples=length(Samplename)
)

sp = sum.pt %>% group_by(Set) %>% summarize(
  med.endo=median(n.endoscopy),
  sd.endo=sd(n.endoscopy),
  med.years=median(n.years),
  med.samples=median(n.samples),
  sd.samples=sd(n.samples)
)

`r pander(as.data.frame(sp))`

```{r echo=F, warning=F, message=F}

#cache.dir = paste(data, 'Analysis', sub('\\+','',paste(tile.w, collapse='_')), sep='/')
model.dir = paste(data, 'Analysis/5e6_arms_all_exAHM0320', sep='/')
split.dir = '/Volumes/research1/gerstung/skillcoyne/Ellie/Analysis/5e6_arms_splits_exAHM0320'

@@

file = paste(model.dir, 'insilico_splits.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
  rocs = insilico.rocs
  performance = insilico.perf
} else {

  split.dirs = list.files(split.dir, full.names = T)
  for (split in 1:length(split.dirs)) {
    print(split)  
    load(list.files(split.dirs[split],full.names = T, pattern='all.*Rdata'), verbose=T)
    performance = rbind(performance, performance.at.1se[[select.alpha]])
    rm(plots,coefs,performance.at.1se,models,cvs)

@@

    preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
    roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
    #rocs = rbind(rocs,c(ci.auc(roc), coords(roc, 'best') ))
    roc$model = 'in-silico'
    roc$is = split
    rocs[[split]] = roc

@@

  insilico.rocs=rocs
  insilico.perf=performance
  save(insilico.rocs, insilico.perf, file=paste(model.dir, 'insilico_splits.Rdata', sep='/'))
}

@@

load(list.files(model.dir, full.names = T, pattern = 'all.*Rdata'))

@@

load(list.files(model.dir,full.names = T, pattern='loo.Rdata'))
rm(plots, performance.at.1se,coefs,nzcoefs,fits)

@@

preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
roc$model = 'discovery'
roc$is = 'discovery'

@@

```{r echo=F, fig.align='left'}
performance$name = 'insilico'
performance$name[which(performance$split == 'discovery')] = 'discovery'
m = melt(performance, measure.vars = 'mean')
ggplot(m, aes(variable,value)) + ylim(0,1) +
  geom_boxplot(fill='mintcream', color='grey55') + geom_jitter(width=0.2, aes(color=name), size=1.5) + 
  geom_text(data=subset(m, value %in% range(value)), aes(label=round(value,2)), vjust=0.02, hjust=0.2) +
  plot.theme + labs(x='CV Models',y='Mean model performance',title='Mean performance for CV model', subtitle='20 insilico 20:69 splits')
```

@@

```
Median AUC `r median(rocs$auc)`, SD `r sd(rocs$auc)`

>>>>

data = '~/Data/Ellie'

@@

@@

sum.pt = summarise.patient.info(all.patient.info)

cache.dir = paste(data, 'Analysis/5e6_arms_all_logR', sep='/')
split.dir = paste(data, 'Analysis/5e6_arms_splits_logR', sep='/')

@@

file = paste(cache.dir, 'insilico_splits.Rdata', sep='/')
if (file.exists(file)) {
  load(file, verbose=T)
  rocs = insilico.rocs
  performance = insilico.perf
} else {
  split.dirs = list.files(split.dir, full.names = T)
  for (split in 1:length(split.dirs)) {
    print(split)  
    models = list.files(split.dirs[split],full.names = T, pattern='all.*Rdata')
    loo = list.files(split.dirs[split],full.names = T, pattern='loo.*Rdata')

    if (length(models) < 1 || length(loo) < 1) {
      warning(paste("Missing model or LOO Rdata file in split", split.dirs[split]))
      next
    }

    load(list.files(split.dirs[split],full.names = T, pattern='all.*Rdata'), verbose=T)
    performance = rbind(performance, performance.at.1se[[select.alpha]])
    rm(plots,coefs,performance.at.1se,models,cvs)

@@

    preds = do.call(rbind.data.frame, lapply(pg.samp, function(df) df[c('Status','Prediction')]))
    roc = pROC::roc(Status ~ Prediction, data=preds, auc=T, ci=T, of='thresholds')
    #rocs = rbind(rocs,c(ci.auc(roc), coords(roc, 'best') ))
    roc$model = 'in-silico'
    roc$is = split
    rocs[[as.character(basename(split.dirs[split]))]] = roc

@@

  insilico.rocs=rocs
  insilico.perf=performance
  save(insilico.rocs, insilico.perf, file=paste(cache.dir, 'insilico_splits.Rdata', sep='/'))
}

@@

load(list.files(cache.dir, full.names = T, pattern = 'all.*Rdata'))

@@

load(list.files(cache.dir,full.names = T, pattern='loo.Rdata'))
rm(plots, performance.at.1se,coefs,nzcoefs,fits)

@@

roc = pROC::roc(Status ~ Prediction, data=predictions, auc=T, ci=T, of='thresholds')
roc$model = 'discovery'
roc$is = 'discovery'

@@

```{r echo=F, fig.align='left'}
performance$name = 'insilico'
performance$name[which(performance$split == 'discovery')] = 'discovery'
m = melt(performance, measure.vars = 'mean')
ggplot(m, aes(variable,value)) + ylim(0,1) +
  geom_boxplot(fill='mintcream', color='grey55', outlier.color = NA) + geom_jitter(width=0.2, aes(color=name), size=1.5) + 
  geom_text(data=subset(m, value %in% range(value)), aes(label=round(value,2)), vjust=0.02, hjust=0.2) +
  plot.theme + labs(x='CV Models',y='Mean model performance',title='Mean performance for CV model', subtitle='20 insilico 20:69 splits')
```

@@

```
Median AUC `r round(median(sapply(rocs, function(x) x$auc)), 2)`, SD `r round(sd(sapply(rocs, function(x) x$auc)), 2)`
